### 2017-08-01 category-theory-basics.tex
Partial proof for Eilenberg-Moore of the comonad being symmetric

\begin{definition}
\label{def:mc}
  A \textbf{monoidal category} $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ is a category $\cat{M}$
  consists of
  \begin{itemize}
  \item a bifunctor $\otimes:\cat{M}\times\cat{M}\rightarrow\cat{M}$, called the tensor product;
  \item an object $I$, called the unit object;
  \item three natural isomorphisms $\alpha$, $\lambda$, and $\rho$ with components
        $$\alpha_{A,B,C}:(A\otimes B)\otimes C\rightarrow A\otimes(B\otimes C)$$
        $$\lambda_A:I\otimes A\rightarrow A$$
        $$\rho_A:A\otimes I\rightarrow A$$
        where $\alpha$ is called associator, $\lambda$ is left unitor, and $\rho$ is right
        unitor,
  \end{itemize}
  such that the following diagrams commute for any objects $A$, $B$, $C$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \square/`->`->`->/<2100,400>[
      ((A\otimes B)\otimes C)\otimes D`
      A\otimes((B\otimes C)\otimes D)`
      (A\otimes B)\otimes(C\otimes D)`
      A\otimes(B\otimes(C\otimes D));
      `
      \alpha_{A\otimes B,C,D}`
      id_A\otimes\alpha_{B,C,D}`
      \alpha_{A,B,C\otimes D}]
    \morphism(0,400)<1100,0>[
      ((A\otimes B)\otimes C)\otimes D`
      (A\otimes(B\otimes C))\otimes D;
      \alpha_{A,B,C}\otimes id_D]
    \morphism(1100,400)<1000,0>[
      (A\otimes(B\otimes C))\otimes D`
      A\otimes((B\otimes C)\otimes D);
      \alpha_{A,B\otimes C,D}]
  \efig
  \and
  \bfig
    \Vtriangle<400,400>[
      (A\otimes I)\otimes B`
      A\otimes(I\otimes B)`
      A\otimes B;
      \alpha_{A,I,B}`
      \rho_A\otimes id_B`
      id_A\otimes\lambda_B]
  \efig
  \end{mathpar}
\end{definition}

\begin{definition}
  A \textbf{Lambek category} (or a \textbf{biclosed monoidal category}) is a monoidal category
  $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ equipped with two bifunctors
  $\rightharpoonup:\cat{M}^{op}\times\cat{M}\rightarrow\cat{M}$ and
  $\leftharpoonup:\cat{M}\times\cat{M}^{op}\rightarrow\cat{M}$ that are both right adjoint to
  the tensor product. That is, the following natural bijections hold:
  \begin{center}
  \begin{math}
  \begin{array}{lll}
    \Hom{L}{X\otimes A}{B}\cong\Hom{L}{X}{A\lto B} & \quad\quad\quad\quad & 
    \Hom{L}{A\otimes X}{B}\cong\Hom{L}{X}{B\rto A}
  \end{array}
  \end{math}
  \end{center}
\end{definition}

\begin{definition}
  A \textbf{symmetric monoidal category} (SMCC) is a monoidal category
  $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ together with a natural transformation with
  components $\e{A,B}:A\otimes B\rightarrow B\otimes A$, called \textbf{exchange}, such that the
  following diagrams commute:
  \begin{mathpar}
  \bfig
    \Vtriangle<300,400>[A\otimes I`I\otimes A`A;\e{A,I}`\rho_A`\lambda_A]
  \efig
  \and
  \bfig
    \Vtriangle/=`->`<-/<300,400>[
      A\otimes B`A\otimes B`B\otimes A;
      id_{A\otimes B}`\e{A,B}`\e{B,A}]
  \efig
  \and
  \bfig
    \hSquares/->`->`->``->`->`->/<400>[
      (A\otimes B)\otimes C`A\otimes(B\otimes C)`(B\otimes C)\otimes A`
      (B\otimes A)\otimes C`B\otimes(A\otimes C)`B\otimes(C\otimes A);
      \alpha_{A,B,C}`\e{A,B\otimes C}`\e{A,B}\otimes id_C``
      \alpha_{B,A,C}`\alpha_{B,A,C}`id_B\otimes\e{A,C}]
  \efig
  \end{mathpar}
\end{definition}

@@

\begin{definition}
  Let $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ and
  $(\cat{M'},\otimes',I',\alpha',\lambda',\rho')$ be monoidal categories. A \textbf{monoidal
  functor} $(F,\m{})$ from $\cat{M}$ to $\cat{M'}$ is a functor $F:\cat{M}\rightarrow\cat{M'}$
  together with a morphism $\m{I}:I'\rightarrow F(I)$ and a natural transformation
  $\m{A,B}:FA'\otimes FB'\rightarrow F(A\otimes B)$, such that the following diagrams commute
  for any objects $A$, $B$, and $C$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \hSquares/->`->`->``->`->`->/<400>[
      (FA\otimes'FB)\otimes'FC`FA\otimes'(FB\otimes'FC)`FA\otimes'F(B\otimes C)`
      F(A\otimes B)\otimes'FC`F((A\otimes B)\otimes C)`F(A\otimes(B\otimes C));
      \alpha'_{FA,FB,FC}`id_{FA}\otimes'\m{A,B}`\m{A,B}\otimes'id_{FC}``
      \m{A,B\otimes C}`\m{A\otimes B,C}`F\alpha_{A,B,C}]
  \efig
  \and
  \bfig
    \square/->`->`<-`->/<600,400>[
      I'\otimes'FA`FA`FI\otimes'FA`F(I\otimes A);
      \lambda'_{FA}`\m{I}\otimes id_{FA}`F\lambda_A`\m{I,A}]
  \efig
  \and
  \bfig
    \square/->`->`<-`->/<600,400>[
      FA\otimes'I'`FA`FA\otimes'FI`F(A\otimes I);
      \rho'_{FA}`\id_{FA}\otimes\m{I}`F\rho_A`\m{A,I}]
  \efig
  \end{mathpar}
\end{definition}

@@

\begin{definition}
  Let $(F,\m{})$ and $(G,\n{})$ be monoidal functors from a monoidal category $\cat{M}$ to a
  monoidal category $\cat{M'}$. A \textbf{monoidal natural transformation} from $(F,\m{})$ to 
  $(G,\n{})$ is a natural transformation $\theta:(F,\m{})\rightarrow(G,\n{})$ such that the
  following diagrams commute for any objects $A$ and $B$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \square<700,400>[
      FA\otimes'FB`F(A\otimes B)`GA\otimes'GB`G(A\otimes B);
      \m{A,B}`\theta_A\otimes'\theta_B`\theta_{A\otimes B}`\n{A,B}]
  \efig
  \and
  \bfig
    \Vtriangle/->`<-`<-/<400,400>[FI`GI`I';\theta_I`\m{I}`\n{I}]
  \efig
  \end{mathpar}
\end{definition}

\begin{definition}
  Let $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ and
  $(\cat{M'},\otimes',I',\alpha',\lambda',\rho')$ be monoidal categories,
  $F:\cat{M}\rightarrow\cat{M'}$ and $G:\cat{M}'\rightarrow\cat{M}$ be functors. The adjunction
  $F:\cat{M}\dashv\cat{M'}:G$ is a \textbf{monoidal adjunction} if $F$ and $G$ are monoidal
  functors, and the unit $\eta$ and the counit $\varepsilon$ are monoidal natural
  transformations.
\end{definition}

@@

\begin{definition}
  \label{def:strong-monad}
  Let $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ be a monoidal category and $(T,\eta,\mu)$ be a
  monad on $\cat{M}$. $T$ is a \textbf{strong monad} if there is natural transformation $\tau$, 
  called the \textbf{tensorial strength}, with components
  $\tau_{A,B}:A\otimes TB\rightarrow T(A\otimes B)$ such that the following diagrams commute:
  \begin{mathpar}
  \bfig
    \Vtriangle<400,400>[I\otimes TA`T(I\otimes A)`TA;\tau_{I,A}`\lambda_{TA}`T\lambda_A]
  \efig
  \and
  \bfig
    \Vtriangle<400,400>[
      A\otimes B`A\otimes TB`T(A\otimes B);id_A\otimes\eta_B`\eta_{A\otimes B}`\tau_{A,B}]
  \efig
  \and
  \bfig
    \square/->`->`->`/<1800,400>[
      (A\otimes B)\otimes TC`T((A\otimes B)\otimes C)`
      A\otimes(B\otimes TC)`T(A\otimes(B\otimes C));
      \tau_{A\otimes B,C}`\alpha_{A,B,TC}`T\alpha_{A,B,C}`]
    \morphism<900,0>[A\otimes(B\otimes TC)`A\otimes T(B\otimes C);id_A\otimes\tau_{B,C}]
    \morphism(900,0)<900,0>[A\otimes T(B\otimes C)`T(A\otimes(B\otimes C));\tau_{A,B\otimes C}]
  \efig
  \and
  \bfig
    \square/`->`->`->/<1400,400>[
      A\otimes T^2B`T^2(A\otimes B)`A\otimes TB`T(A\otimes B);
      `id_A\otimes\mu_B`\mu_{A\otimes B}`\tau_{A,B}]
    \morphism(0,400)<700,0>[A\otimes T^2B`T(A\otimes TB);\tau_{A,TB}]
    \morphism(700,400)<700,0>[T(A\otimes TB)`T^2(A\otimes B);T\tau_{A,B}]
  \efig
  \end{mathpar}
\end{definition}

>>>>

\begin{definition}
\label{def:mc}
  A \textbf{monoidal category} $(\cat{M},\tri,I,\alpha,\lambda,\rho)$ is a category $\cat{M}$
  consists of
  \begin{itemize}
  \item a bifunctor $\tri:\cat{M}\times\cat{M}\rightarrow\cat{M}$, called the tensor product;
  \item an object $I$, called the unit object;
  \item three natural isomorphisms $\alpha$, $\lambda$, and $\rho$ with components
        $$\alpha_{A,B,C}:(A\tri B)\tri C\rightarrow A\tri(B\tri C)$$
        $$\lambda_A:I\tri A\rightarrow A$$
        $$\rho_A:A\tri I\rightarrow A$$
        where $\alpha$ is called associator, $\lambda$ is left unitor, and $\rho$ is right
        unitor,
  \end{itemize}
  such that the following diagrams commute for any objects $A$, $B$, $C$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \square/`->`->`->/<2100,400>[
      ((A\tri B)\tri C)\tri D`
      A\tri((B\tri C)\tri D)`
      (A\tri B)\tri(C\tri D)`
      A\tri(B\tri(C\tri D));
      `
      \alpha_{A\tri B,C,D}`
      id_A\tri\alpha_{B,C,D}`
      \alpha_{A,B,C\tri D}]
    \morphism(0,400)<1100,0>[
      ((A\tri B)\tri C)\tri D`
      (A\tri(B\tri C))\tri D;
      \alpha_{A,B,C}\tri id_D]
    \morphism(1100,400)<1000,0>[
      (A\tri(B\tri C))\tri D`
      A\tri((B\tri C)\tri D);
      \alpha_{A,B\tri C,D}]
  \efig
  \and
  \bfig
    \Vtriangle<400,400>[
      (A\tri I)\tri B`
      A\tri(I\tri B)`
      A\tri B;
      \alpha_{A,I,B}`
      \rho_A\tri id_B`
      id_A\tri\lambda_B]
  \efig
  \end{mathpar}
\end{definition}

\begin{definition}
  A \textbf{Lambek category} (or a \textbf{biclosed monoidal category}) is a monoidal category
  $(\cat{M},\tri,I,\alpha,\lambda,\rho)$ equipped with two bifunctors
  $\rightharpoonup:\cat{M}^{op}\times\cat{M}\rightarrow\cat{M}$ and
  $\leftharpoonup:\cat{M}\times\cat{M}^{op}\rightarrow\cat{M}$ that are both right adjoint to
  the tensor product. That is, the following natural bijections hold:
  \begin{center}
  \begin{math}
  \begin{array}{lll}
    \Hom{L}{X\tri A}{B}\cong\Hom{L}{X}{A\lto B} & \quad\quad\quad\quad & 
    \Hom{L}{A\tri X}{B}\cong\Hom{L}{X}{B\rto A}
  \end{array}
  \end{math}
  \end{center}
\end{definition}

\begin{definition}
  \label{def:smcc}
  A \textbf{symmetric monoidal category} (SMCC) is a monoidal category
  $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ together with a natural transformation with
  components $\e{A,B}:A\otimes B\rightarrow B\otimes A$, called \textbf{exchange}, such that the
  following diagrams commute:
  \begin{mathpar}
  \bfig
    \Vtriangle<300,400>[A\otimes I`I\otimes A`A;\e{A,I}`\rho_A`\lambda_A]
  \efig
  \and
  \bfig
    \Vtriangle/=`->`<-/<300,400>[
      A\otimes B`A\otimes B`B\otimes A;
      id_{A\otimes B}`\e{A,B}`\e{B,A}]
  \efig
  \and
  \bfig
    \hSquares/->`->`->``->`->`->/<400>[
      (A\otimes B)\otimes C`A\otimes(B\otimes C)`(B\otimes C)\otimes A`
      (B\otimes A)\otimes C`B\otimes(A\otimes C)`B\otimes(C\otimes A);
      \alpha_{A,B,C}`\e{A,B\otimes C}`\e{A,B}\otimes id_C``
      \alpha_{B,A,C}`\alpha_{B,A,C}`id_B\otimes\e{A,C}]
  \efig
  \end{mathpar}
\end{definition}

We use $\tri$ for non-symmetric monoidal categories while $\otimes$ for symmetric ones.

@@

\begin{definition}
  Let $(\cat{M},\tri,I,\alpha,\lambda,\rho)$ and
  $(\cat{M'},\tri',I',\alpha',\lambda',\rho')$ be monoidal categories. A \textbf{monoidal
  functor} $(F,\m{})$ from $\cat{M}$ to $\cat{M'}$ is a functor $F:\cat{M}\rightarrow\cat{M'}$
  together with a morphism $\m{I}:I'\rightarrow F(I)$ and a natural transformation
  $\m{A,B}:FA'\tri FB'\rightarrow F(A\tri B)$, such that the following diagrams commute
  for any objects $A$, $B$, and $C$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \hSquares/->`->`->``->`->`->/<400>[
      (FA\tri'FB)\tri'FC`FA\tri'(FB\tri'FC)`FA\tri'F(B\tri C)`
      F(A\tri B)\tri'FC`F((A\tri B)\tri C)`F(A\tri(B\tri C));
      \alpha'_{FA,FB,FC}`id_{FA}\tri'\m{A,B}`\m{A,B}\tri'id_{FC}``
      \m{A,B\tri C}`\m{A\tri B,C}`F\alpha_{A,B,C}]
  \efig
  \and
  \bfig
    \square/->`->`<-`->/<600,400>[
      I'\tri'FA`FA`FI\tri'FA`F(I\tri A);
      \lambda'_{FA}`\m{I}\tri id_{FA}`F\lambda_A`\m{I,A}]
  \efig
  \and
  \bfig
    \square/->`->`<-`->/<600,400>[
      FA\tri'I'`FA`FA\tri'FI`F(A\tri I);
      \rho'_{FA}`\id_{FA}\tri\m{I}`F\rho_A`\m{A,I}]
  \efig
  \end{mathpar}
\end{definition}

@@

\begin{definition}
  Let $(F,\m{})$ and $(G,\n{})$ be monoidal functors from a monoidal category
  $(\cat{M},\otimes,I,\alpha,\lambda,\rho)$ to a monoidal category
  $(\cat{M'},\otimes',I',\alpha',\lambda',\rho')$. A \textbf{monoidal natural transformation}
  from $(F,\m{})$ to $(G,\n{})$ is a natural transformation $\theta:(F,\m{})\rightarrow(G,\n{})$
  such that the following diagrams commute for any objects $A$ and $B$ in $\cat{M}$:
  \begin{mathpar}
  \bfig
    \square<700,400>[
      FA\tri'FB`F(A\tri B)`GA\tri'GB`G(A\tri B);
      \m{A,B}`\theta_A\tri'\theta_B`\theta_{A\tri B}`\n{A,B}]
  \efig
  \and
  \bfig
    \Vtriangle/->`<-`<-/<400,400>[FI`GI`I';\theta_I`\m{I}`\n{I}]
  \efig
  \end{mathpar}
\end{definition}

\begin{definition}
  Let $(\cat{M},\tri,I,\alpha,\lambda,\rho)$ and
  $(\cat{M'},\tri',I',\alpha',\lambda',\rho')$ be monoidal categories,
  $F:\cat{M}\rightarrow\cat{M'}$ and $G:\cat{M}'\rightarrow\cat{M}$ be functors. The adjunction
  $F:\cat{M}\dashv\cat{M'}:G$ is a \textbf{monoidal adjunction} if $F$ and $G$ are monoidal
  functors, and the unit $\eta$ and the counit $\varepsilon$ are monoidal natural
  transformations.
\end{definition}

@@

\begin{definition}
  \label{def:strong-monad}
  Let $(\cat{M},\tri,I,\alpha,\lambda,\rho)$ be a monoidal category and $(T,\eta,\mu)$ be a
  monad on $\cat{M}$. $T$ is a \textbf{strong monad} if there is natural transformation $\tau$, 
  called the \textbf{tensorial strength}, with components
  $\tau_{A,B}:A\tri TB\rightarrow T(A\tri B)$ such that the following diagrams commute:
  \begin{mathpar}
  \bfig
    \Vtriangle<400,400>[I\tri TA`T(I\tri A)`TA;\tau_{I,A}`\lambda_{TA}`T\lambda_A]
  \efig
  \and
  \bfig
    \Vtriangle<400,400>[
      A\tri B`A\tri TB`T(A\tri B);id_A\tri\eta_B`\eta_{A\tri B}`\tau_{A,B}]
  \efig
  \and
  \bfig
    \square/->`->`->`/<1800,400>[
      (A\tri B)\tri TC`T((A\tri B)\tri C)`
      A\tri(B\tri TC)`T(A\tri(B\tri C));
      \tau_{A\tri B,C}`\alpha_{A,B,TC}`T\alpha_{A,B,C}`]
    \morphism<900,0>[A\tri(B\tri TC)`A\tri T(B\tri C);id_A\tri\tau_{B,C}]
    \morphism(900,0)<900,0>[A\tri T(B\tri C)`T(A\tri(B\tri C));\tau_{A,B\tri C}]
  \efig
  \and
  \bfig
    \square/`->`->`->/<1400,400>[
      A\tri T^2B`T^2(A\tri B)`A\tri TB`T(A\tri B);
      `id_A\tri\mu_B`\mu_{A\tri B}`\tau_{A,B}]
    \morphism(0,400)<700,0>[A\tri T^2B`T(A\tri TB);\tau_{A,TB}]
    \morphism(700,400)<700,0>[T(A\tri TB)`T^2(A\tri B);T\tau_{A,B}]
  \efig
  \end{mathpar}
\end{definition}
